### 2023-08-21 paper/coverletter/nhb1.tex
rejected

\documentclass{stanfordletter}
%\makelabels
\begin{document}
	\name{Veronica Boyce}
	\signature{Veronica Boyce \\ Maya Mathur \\ Michael C. Frank}

	\begin{letter}{Dr. Stavroula Kousta \\ Chief Editor \\ Nature Human Behavior}

          \opening{Dear Dr. Kousta,} 
          This
          submission contains the manuscript ``Eleven years of student replication projects provide evidence on
          the correlates of replicability in psychology''
          which we are submitting to Nature Human Behavior for
          publication.

          The manuscript reports on a new dataset of 176 replications of social science experiments conducted by students in a graduate-level methods course. The roughly 170 existing replications from RP:P, ManyLabs, and Camerer et al. (2018) have been heavily discussed and re-analysed. Our current dataset is an independent sample of replications that greatly increases the number of replications.  In the paper, we discuss how being able to replicate a study is a prerequisite to building cumulative science on that finding. Student replications were successful 49\% of the time, indicating that many published findings are out of reach of student investigators.    We examine correlates of replication success, including   statistical, demographic, and  experimental design features.  We find that studies with large original effect sizes and within-subjects designs were the most likely to replicate.

          We believe this paper will be of broad
          interest to psychologists.

          Researchers with relevant expertise and interest who
          therefore might be appropriate reviewers include Brian Nosek, Simine Vazire, Dani\"{e}l Lakens, and Daniel Quintana.

          This manuscript is not under review at any other
          venue. Please let us know if you need any further
          information in connection with this submission. We look
          forward to hearing from you regarding the manuscript!
          \closing{Sincerely,}

	\end{letter}

\end{document}